Language Skills in an Ethnically Segmented Labour market: Estonia 1989 -- 2012

### 2018-07-22 paper/ijm.tex
Minor corrections

\citet{leslie+lindley2001} analyse both employment and income but
focus on a single year and English language only.
\citet{YaoandOurs2015} only analyse the role of Dutch language
on labour market performance in Netherlands.  
\fxwarning{ot:add CEE stuff here?}
\fxwarning{SKB: Let's collect first enough material and then decide, but I can start adding here some notes from \cite{Kahanec2009,Kahanec2010}.}
%Start here CEE stuff for now

@@

The Estonian economy, a part of the Soviet economy up to 1991, successfully
re-oriented to the Western markets and has been growing rapidly since
the mid-1990s.  \Cref{fig:gdp_growth} indicates that the Estonian economy grew around 7\% per year before the financial crisis in 2008 and after that grew at a lower rate around 3\% per year.
%\fxwarning{ot: a few numbers about economic growth would be good}
%\fxwarning{SKB: I added rather a figure. Pure numbers or table might not give the best impression of the growth.}
\begin{figure}[h]
	\centering                      
	\caption{GDP growth per year for Estonia}\label{fig:gdp_growth}
	\includegraphics[width=0.6\linewidth]{gdp_growth_per_year.eps}
	\caption*{Source: Worldbank - World Development Indicators, Series  NY.GDP.MKTP.KD.ZG}
\fxnote{ot: remove the cryptic series?}
%\fxnote{skb: removed the series from y-axis label. Should I remove the rest as well since it is already stated in the figure caption?}
\end{figure}
e-adjustment, English skills become extremely
valuable while the supply was still rather limited during the early
years of transition.

@@

The central results for unemployment regressions are given in
\cref{tab:unemployment_estimation_by_sex_and_ethnic_same_sample}.\footnote{We
  restrict both models to have the same number of observations as
  requested by a reviewer. The unrestricted table is in the appendix \cref{tab:unemployment_estimation_by_sex_and_ethnic}.}
The
table clearly indicates that Estonian language skills are associated
with approximately 5 percentage points lower unemployment for both
Russian men and women. We also see that the effect is not radically
influenced by the individual characteristics--for men it is
virtually unchanged while for women the estimate falls by 30\%, from
6.5 to 4.5 percentage points.

@@

We repeat the exercise for unemployment models.
However, as most of the results are not statistically significant, we
do not report the results here.

@@

The most surprising result in our analysis so far--the fact that
Russian men earn virtually no Estonian language premium while women do--warrants a
deeper look at the data.  We split the average wage for the Russian male wage-earners
along the three
dimensions: Estonian language skills (\emph{EST 0} and \emph{EST 1}),
residence in Tallinn metro area (\emph{Tallinn} and \emph{Other}),
and education (\emph{< HS}, \emph{HS}, \emph{College Degree}).  The metropolitan area is one of the two major regions
where the Russian-speaking population is concentrated, and it's
booming economy may be very different from the other such region, the
North-East (Ida-Virumaa), and the rest of the country.
Even more, while Russians form a large minority in and around Tallinn,
they are in majority in the North-East.  
In order to correct for the strong wage and language skill trends
through the observation period, we do not present the average wage,
but instead the respective OLS residuals where we explain the wage by year fixed
effects and include no other explanatory variables (Table~\ref{tab:loc-lang-edu}).

@@

We can conclude from the table~\ref{tab:loc-lang-edu} that our
most surprising finding, no positive effect of Estonian language
skills for Russian men, is primarily driven by workers with
high school degree.  It also suggests that highly educated workers in
the metropolitan area enjoy, in fact, a substantial language premium.
However, this group is not very large.

@@

The other results are more in line with other studies.
English language is related to a substantial income premium but
virtually to no effect on unemployment.  This is similar to findings by
\citet{ginsburgh+prieto-rodriguez2011ILRR} and
\citet{beblavy+2016CEPS} who show that English is more important at better paid
jobs.  In contrast, our outcomes suggest that English has little use
at unstable jobs, at these jobs where workers frequently move between
employment and unemployment.
Note also that for
English, we are able to explain a substantial part of the raw effect
by individual characteristics, suggesting that the true causal effect
may be smaller than our estimates.

>>>>

\citet{leslie+lindley2001} analyse both employment and income but
focus on a single year and English language only.
\citet{YaoandOurs2015} only analyse the role of Dutch language
on labour market performance in Netherlands.  
\fxwarning{ot:add CEE stuff here?}
\fxwarning{SKB: Let's collect first enough material and then decide, but I can start adding here some notes from \cite{Kahanec2009,Kahanec2010}.}
%Start here CEE stuff for now

@@

The Estonian economy, a part of the Soviet economy up to 1991, successfully
re-oriented to the Western markets and has been growing rapidly since
the mid-1990s.  
\Cref{fig:gdp_growth} indicates that the Estonian economy grew around 7\% per year before the financial crisis in 2008 and after that grew at a lower rate around 3\% per year.
%\fxwarning{ot: a few numbers about economic growth would be good}
%\fxwarning{SKB: I added rather a figure. Pure numbers or table might not give the best impression of the growth.}
\begin{figure}[h]
	\centering                      
	\caption{GDP growth per year for Estonia}\label{fig:gdp_growth}
	\includegraphics[width=0.6\linewidth]{gdp_growth_per_year.eps}
	\caption*{Source: Worldbank - World Development Indicators, Series  NY.GDP.MKTP.KD.ZG}
\fxnote{ot: remove the cryptic series?}
%\fxnote{skb: removed the series from y-axis label. Should I remove the rest as well since it is already stated in the figure caption?}
\end{figure}
 As a bi-product of the re-adjustment, English skills become extremely
valuable while the supply was still rather limited during the early
years of transition.

@@

The central results for unemployment regressions are given in
\cref{tab:unemployment_estimation_by_sex_and_ethnic_same_sample}\footnote{We
  restrict both models to have the same number of observations as
  requested by a reviewer. The unrestricted table is in the appendix \cref{tab:unemployment_estimation_by_sex_and_ethnic}.}.
The
table clearly indicates that Estonian language skills are associated
with approximately 5 percentage points lower unemployment for both
Russian men and women. We also see that the effect is not radically
influenced by the individual characteristics--for men it is
virtually unchanged while for women the estimate falls by 30\%, from
6.5 to 4.5 percentage points.

@@

We repeat the exercise for the unemployment models.
However, as most of the results are not statistically significant, we
do not report the results here.

@@

The most surprising result in our analysis so far -- the fact that
Russian men earn virtually no Estonian language premium while women do -- warrants a
deeper look at the data.  We split the average wage for the Russian male wage-earners
along the three
dimensions: Estonian language skills (\emph{EST 0} and \emph{EST 1}),
residence in Tallinn metro area (\emph{Tallinn} and \emph{Other}),
and education (\emph{< HS}, \emph{HS}, \emph{College Degree}).  The metropolitan area is one of the two major regions
where the Russian-speaking population is concentrated, and it's
booming economy may be very different from the other such region, the
North-East (Ida-Virumaa), and the rest of the country.
Even more, while Russians form a large minority in and around Tallinn,
they are in majority in the North-East.  
In order to correct for the strong wage and language skill trends
through the observation period, we do not present the average wage,
but instead the respective OLS residuals where we explain the wage by year fixed
effects and include no other explanatory variables (Table~\ref{tab:loc-lang-edu}).

@@

We can conclude from the Table~\ref{tab:loc-lang-edu} that our
most surprising finding, no positive effect of Estonian language
skills for Russian men, is primarily driven by workers with
high school degree.  It also suggests that highly educated workers in
the metropolitan area enjoy, in fact, a substantial language premium.
However, this group is not very large.

@@

The other results are more in line with other studies.
English language is related to a substantial income premium but
virtually to no effect on unemployment.  This is similar to findings by
\citet{ginsburgh+prieto-rodriguez2011ILRR} and \cite{fabo+2017E}
 who show that English is more important at better paid
jobs.  In contrast, our outcomes suggest that English has little use
at unstable jobs, at these jobs where workers frequently move between
employment and unemployment.
Note also that for
English, we are able to explain a substantial part of the raw effect
by individual characteristics, suggesting that the true causal effect
may be smaller than our estimates.
